---
title: "Mo_E18_5_MesCells"
author: "Irepan Salvador-Martinez"
date: "30/09/2021"
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, out.width = "100%", fig.align = "center",
  message = FALSE, warning = FALSE
)
options(width = 900)
```


# Mo_E18_5_MesCells Samples

We did scRNAseq (10X version 3) of mouse endochondral bones cell populations (in a balanced/even representation way, by FACS isolation) in perinatal stages E18 (just before birth) and postnatal day 1 (PN1).

## About input files

The files used as input for this analysis are the files coming out of the the
*cellranger count* pipeline.

In this case I used cellranger v6.0.0.



```{r packages}
library(Matrix)
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(DoubletFinder)
library(BiocStyle)
library(dittoSeq)
library(RColorBrewer)
```

```{r}

ggplotColours <- function(n = 6, h = c(0, 360) + 15){
  if ((diff(h) %% 360) < 1) h[2] <- h[2] - 360/n
  hcl(h = (seq(h[1], h[2], length = n)), c = 100, l = 65)
}

```


## Loading data from 10X multi-modal experiments

Seurat is also able to analyze data from multimodal 10X experiments processed using CellRanger v3.

I first read all the data and then create a seurat object with the gene expression, and append the antibody and multiplex assays

```{r Load precomputed data}
Mo_E18_5_MesCells <-  readRDS("Mo_E18_5_MesCells_preProcess.rds")
# define assay
DefaultAssay(Mo_E18_5_MesCells) <- "RNA"
Mo_E18_5_MesCells_nonblood <- readRDS("Mo_E18_5_MesCells_nonblood.rds")
DefaultAssay(Mo_E18_5_MesCells_nonblood) <- "RNA"
```


```{r Read 10X data,eval=FALSE}
# Load the Mo_E18_5_MesCells dataset
cluster_dir <- "~/scratch/stromal_cells/1-cellranger_mapping/Mo_E18_5_MesCells/Mo_E18_5_MesCells"
Mo_E18_5_MesCells.data <- Read10X(data.dir = paste0(cluster_dir,"/outs/filtered_feature_bc_matrix/"))
# Initialize the Seurat object with the raw (non-normalized data).
Mo_E18_5_MesCells <- CreateSeuratObject(counts = Mo_E18_5_MesCells.data, 
                                        project = "Mo_E18_5_MesCells",
                                        min.cells = 3, min.features = 200)
rm(Mo_E18_5_MesCells.data)
saveRDS(Mo_E18_5_MesCells, "Mo_E18_5_MesCells_init.rds")
```

```{r}
Mo_E18_5_MesCells <- readRDS("Mo_E18_5_MesCells_init.rds")
```

