---
title: "E18 and PN1 CC_Phase and Doublets removal"
author: "Irepan Salvador-Martinez"
date: "07/12/2021"
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  out.width = "100%", fig.align = "center",fig.width = 8,fig.height = 5,
  message = FALSE, warning = FALSE
)
options(width = 900)
```



# About this analysis

After the integration of the samples, I will remove the doublets and will compute
the Cell cycle phase score.

After this I will export the datasets for velocyto analysis.


```{r packages}
library(dplyr)
library(Seurat)
library(cowplot)
library(patchwork)
library(ggplot2)
library(BiocStyle)
library(harmony)
library(dittoSeq)
library(RColorBrewer)

library(SeuratDisk)
library(SeuratData)
```
# Preprocess

## Load pre-processed data



```{r}
E18_PN1_Harmony <- readRDS("E18_PN1_Harm_Phase_doublets.rds")
```


# Assign pre-defined color palette

```{r}
mylevels <- c("PaS/EndostealFibroblasts","Neutrophils1","Monocytes","Neutrophils2",
              "CAR","Osteoclasts","Chondrogenic","HPC","B_cell","OCP","Eo/Bas","PFP","PHP",
              "Myofibroblasts","Erythroid","Endothelial","Osteogenic","Schwann",
              "Tenogenic","CLP_NK_Tcells","doublets")

E18_PN1_Harmony@meta.data$Cell_Type <- factor(E18_PN1_Harmony@meta.data$Cell_Type, levels = mylevels)
```

```{r}
col_palette <- c( rgb(0, 176/255, 240/255),       # C0=PaS/ Endosteal Fibroblasts (0, 176, 240)
                  rgb(1/255, 203/255,0),          # Neutrophils 1 (1, 203,0)
                  rgb(213/255, 252/255, 121/255), # Monocytes (213, 252,121)
                  rgb(16/255, 152/255, 85/255),   # Neutrophils2 (16, 152, 85)
                  rgb(255/255, 126/255, 72/255),  # CAR (255, 126, 72)
                  rgb(174/255, 205/255, 100/255), # Osteoclasts (174, 205, 100)
                  rgb(1/255, 134/255, 227/255),   # Chondrogenic Lineage (1, 134, 227)
                  rgb(62/255, 243/255, 180/255),  #C7=HPC (hematopoietic progenitor cells) (62, 243, 180)
                  rgb(101/255, 203/255, 176/255), #C8= B cell lineage (101, 203, 176)
                  rgb(0, 145/255, 147/255),       #C9=OCP (0, 145, 147)
                  rgb(122/255, 129/255, 255/255), #C10= Eo/Bas (122, 129, 255)
                  rgb(121/255, 249/255, 255/255), #C11= Proliferating Fibroblasts Progenitors (PFP) (121, 249, 255)
                  rgb(3/255, 50/255, 255/255),    #C12 = Proliferating Hematopoietic Progenitors (PHP) (3, 50, 255)
                  rgb(255/255, 47/255, 146/255),  #C13= Myofibroblasts (255, 47, 146)
                  rgb(205/255, 31/255, 113/255),  #C14= Erythroid cells (205, 31, 113)
                  rgb(255/255, 191/255, 0/255),   #C15= EC (endothelial cells) (255, 191, 0)
                  rgb(110/255, 68/255, 244/255),  #C16= Osteogenic lineage (110, 68, 244)
                  rgb(244/255, 56/255, 234/255),  #C17=Schwann cells(244, 56, 234)
                  rgb(254/255, 138/255, 216/255), #C18= Tenogenic lineage (254, 138, 216)
                  rgb(191/255,144/255, 0),        #C19= CLP NK&Tcell (191,144, 0)
                  rgb(0.6,0.6,0.6))

DimPlot(E18_PN1_Harmony, group.by = "Cell_Type", label = T, cols = col_palette, repel = T )
```


## Visualise by library

```{r, fig.width=12, fig.height=6}
DimPlot(E18_PN1_Harmony, group.by = "Cell_Type", split.by = "orig.ident", pt.size = 0.3,
        label = T, cols = col_palette, repel = T ) + theme_bw() + NoLegend() 

```

```{r, eval=FALSE}
#p1 <- DimPlot(E18_PN1_Harmony, group.by = "Cell_Type", cols = col_palette, label = F ) + NoLegend()
#doublets <- CellSelector(p1)
#write.csv(x = doublets, file = "doublets_to_remove.csv")
d <- read.csv(file = "doublets_to_remove.csv")
doublets <- d$x
allcells <- colnames(E18_PN1_Harmony)
E18_PN1_Harmony <- subset(E18_PN1_Harmony, cells = setdiff(allcells,doublets) )
DimPlot(E18_PN1_Harmony, group.by = "Cell_Type", cols = col_palette, label = F )

```

## Get cell cycle phase

In here I will assing a cellcycle score to the cells to know if they are in 
proliferative (G2M and S phase) or non proliferative (G1) stage.

This is to check if any cluster we find is dominated by cells in a specific cell
cycle stage.

```{r, eval=FALSE}
# Basic function to convert human to mouse gene names
#convertHumanGeneList <- function(x){
#require("biomaRt")
#human = useMart("ensembl", dataset = "hsapiens_gene_ensembl")
#mouse = useMart("ensembl", dataset = "mmusculus_gene_ensembl")
#genesV2 = getLDS(attributes = c("hgnc_symbol"), filters = "hgnc_symbol", values = x , mart = human, attributesL = c("mgi_symbol"), martL = mouse, uniqueRows=T)
#humanx <- unique(genesV2[, 2])
# Print the first 6 genes found to the screen
#print(head(humanx))
#return(humanx)
#}
```

```{r, eval=FALSE}
#m.s.genes <- convertHumanGeneList(cc.genes.updated.2019$s.genes)
#m.g2m.genes <- convertHumanGeneList(cc.genes.updated.2019$g2m.genes)

#write.csv(m.s.genes, file = "m.s.genes.csv")
#write.csv(m.g2m.genes, file = "m.g2m.genes.csv")

```


```{r, eval=FALSE}
m.s.genes  <- read.csv(file = "m.s.genes.csv", header = T, row.names = 1)
m.s.genes  <- m.s.genes[,1]
m.g2m.genes<- read.csv(file = "m.g2m.genes.csv", header = T, row.names = 1)
m.g2m.genes <- m.g2m.genes[,1]
```

```{r, eval=FALSE}
E18_PN1_Harmony <- CellCycleScoring(E18_PN1_Harmony, s.features = m.s.genes, 
                                    g2m.features = m.g2m.genes, set.ident = TRUE)
```



```{r}
DimPlot(E18_PN1_Harmony, group.by = "Phase", label = F,
        cols= c("red", "green", "blue") )
```


## Export for velocyto by library


```{r, eval=FALSE}
########

E18_PN1_Harmony@meta.data$Cell_Type <- as.character(E18_PN1_Harmony@meta.data$Cell_Type)

libs <- SplitObject(E18_PN1_Harmony, split.by = "orig.ident")

for (i in 1:length(libs))
{
  libname <- unique(libs[[i]]@meta.data$orig.ident)
  print(libname)
    ## --------------------
  count_matrix <- libs[[i]]@assays$RNA@counts
  Stromal_scratch <- CreateSeuratObject(counts = count_matrix, project = libname, assay = "RNA",
                                     meta.data = libs[[i]]@meta.data[, c(2:4,15,23)])
  Stromal_scratch@reductions$umap.rna <- libs[[i]]@reductions$umap
  # then we save it as a Seurat h5 object
  #intermediate_h5_file=paste(input_file,".h5Seurat",sep="")
  intermediate_h5_file= paste0("../3.4-RNAvel/h5ad/",libname,"_Rib.h5Seurat")
  SaveH5Seurat(Stromal_scratch, filename = intermediate_h5_file, overwrite = T)
  Convert(intermediate_h5_file, dest = "h5ad",overwrite = T)
}


```


```{r, eval=FALSE}
saveRDS(E18_PN1_Harmony, "E18_PN1_Harm_Phase_doublets.rds")
```


# Export RDS file for viz

```{r, eval=FALSE}
#source(file = "../../utils/seurat2shiny.R")

#seurat2shiny(E18_PN1_Harmony, reduction = "umap", save = TRUE, name = "E18_PN1_Harmony")
```



# Session Info

```{r}
sessionInfo()
```

